---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Initilization of libraries
```{r init, warning = FALSE, message = FALSE}
# using warning = FALSE and message = FALSE to supress messages
library(ggplot2)
library(lattice)
library(dplyr)
```

## Loading and preprocessing data
```{r load data}
data<-read.csv("activity.csv")
data$date<-as.Date(as.character(data$date))

```




## What is mean total number of steps taken per day?
```{r  steps}
data<-group_by(data,date)
total_steps <- summarise(data, sum(steps, na.rm=TRUE))[[2]]
hist(total_steps)
mean_daily_steps<-mean(total_steps)
median_daily_steps<-median(total_steps)
```
Mean daily steps: `r mean_daily_steps`  
Median daily steps: `r median_daily_steps`

## What is the average daily activity pattern?

```{r daily pattern}
data<-group_by(data, interval)
pattern<-summarise(data, mean(steps, na.rm = TRUE))[[2]]
interval<-summarise(data, mean(steps, na.rm = TRUE))[[1]]
plot(interval, pattern, type = "l")
```


##Imputing missing values
```{r missing values}
avg_steps_per_interval<- cbind(pattern, interval)
data2<- merge(data, avg_steps_per_interval)
missing<- is.na(data2$steps)
num_missing<-sum(missing)
data2$steps[missing]<-data2$pattern[missing]
data2<-select(data2, -pattern)
data2<-arrange(data2, date, interval)
data2<-group_by(data2, date)
new_steps<- summarise(data2, sum(steps, na.rm=TRUE))[[2]]
hist(new_steps)
mean_daily_steps2<-mean(new_steps)
median_daily_steps2<-median(new_steps)
```
For any missing value of steps, the average # of steps for that time interval is used.
Number of missing values in "steps": `r num_missing`  
Mean daily steps: `r mean_daily_steps2`   
Median daily steps: `r median_daily_steps2`  

After imputing missing values, mean and median number of total dailiy steps has gone up.  



## Are there differences in activity patterns between weekdays and weekends?

```{r weekend vs weekday}
data2$day<-weekdays(data2$date)
data2$weekend[data2$day == "Sunday"]<-"weekend"
data2$weekend[data2$day == "Monday"]<-"weekday"
data2$weekend[data2$day == "Tuesday"]<-"weekday"
data2$weekend[data2$day == "Wednesday"]<-"weekday"
data2$weekend[data2$day == "Thursday"]<-"weekday"
data2$weekend[data2$day == "Friday"]<-"weekday"
data2$weekend[data2$day == "Saturday"]<-"weekend"
data2$weekend<-as.factor(data2$weekend)
data2<-group_by(data2,weekend, interval )
temp<-summarize(data2, mean(steps, na.rm=TRUE))
weekday<-as.character(temp$weekend) == "weekday"
weekend<-as.character(temp$weekend) == "weekend"
weekday_data<- temp[[3]][weekday]
weekday_data<-temp[[3]][weekend]
names(temp)<- c("weekend", "interval", "number_of_steps")
xyplot(number_of_steps ~ interval | weekend, data=temp, layout =  c(1,2), type = "l")
```
